Think deep for reef resilience

### 2025-08-10 paper/paper_draft.qmd
Add text and references on GCM development families

The five selected GCMs span a range of Equilibrium Climate Sensitivity (ECS) values within
the Intergovernmental Panel on Climate Change (IPCC) assessed likely range, and a range of
GCM development families \[**TI: Ben, what does "GCM development families" mean?**\]. The
IPCC defines ECS as the long term warming caused by a doubling of carbon dioxide above its
pre-industrial concentration [@forster2021]. The five GCMs therefore allow for a range of
possible climate futures to be explored, capturing the uncertainty in future responses to
CO2 increases (@sfig-ecs; @fig-methods-dhws; [@grose2023; @shepherd2018]). Projections using
Shared Socio-economic Pathway (SSP) 2-4.5 was used to represent the plausible emission
pathway. It represents a moderate emissions scenario with continued CO2 emissions comparable
to present day until 2050, after which emissions gradually decline [@hausfather2020; IPCC
-@climate2023]. The selected GCMs represent 2.5°C to 4.31°C of global atmospheric warming in
the 2081-2100 period relative to 1850-1900.

@@

Timeseries clustering analyses were performed at the reefal bioregion scale (defined in
\[https://geohub-gbrmpa.hub.arcgis.com/datasets/GBRMPA::reefal-marine-bioregions/about\]).
Additional analyses were also performed at larger spatial scales, such as the GBRMPA
management regions and whole-of-GBR scale. These were found to lead to identical conclusions
and so are not shown here. Results of these analyses can be found at \[**insert citation for
repository**\] for those interested.

>>>>

The five selected GCMs span a range of Equilibrium Climate Sensitivity (ECS) values within
the Intergovernmental Panel on Climate Change (IPCC) assessed likely range. The IPCC defines
ECS as the long term warming caused by a doubling of carbon dioxide above its pre-industrial
concentration [@forster2021]. GCM development institutions regularly share aspects of model
development. Brunner et al. [-@brunner2020] assessed the independence of 33 models,
quantitatively classifying models into 'model development families' based on patterns in
atmospheric temperature and sea level pressure. By selecting five GCMs for investigation
that span a range of four 'model development families' [@brunner2020] (note: we assume
GFDL-CM4 aligns with GFDL-ESM4; ACCESS-ESM1-5 and ACCESS-CM2 occupy the same family), we are
able to assess climate projections over a broad range of independently derived futures
[@abramowitz2019]. The selected GCMs therefore allow for a range of possible climate futures
to be explored, capturing the uncertainty in future responses to CO2 increases (@sfig-ecs;
@fig-methods-dhws; [@grose2023; @shepherd2018]). Projections using Shared Socio-economic
Pathway (SSP) 2-4.5 was used to represent the plausible emission pathway. It represents a
moderate emissions scenario with continued CO2 emissions comparable to present day until
2050, after which emissions gradually decline [@hausfather2020; IPCC -@climate2023]. The
selected GCMs represent 2.5°C to 4.31°C of global atmospheric warming in the 2081-2100
period relative to 1850-1900.

@@

Timeseries clustering analyses were performed at the reefal bioregion scale (defined in
GBRMPA [-@GBRMPA2001]). Additional analyses were also performed at larger spatial scales,
such as the GBRMPA management regions and whole-of-GBR scale. These were found to lead to
identical conclusions and so are not shown here. Results of these analyses can be found at
\[**insert citation for repository**\] for those interested.
